---
title: "RStudio Datasets"
author: "Nicholas Moore"
date: "2022-07-29 - Present"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## {.tabset .tabset-fade} 

**Background:** Hi! This is my journey of navigating RStudio and becoming an overall more knowledgeable data analyst / scientist. Here I will create a variety of storytelling visuals with RStudio's built-in datasets. 

### Iris  



**Loading Data / Basic Plotting Tutorial**
```{r message=FALSE, warning=FALSE}
library(ggplot2) #qplot (basic scatterplot) 
library(datasets)
library(tidyverse)
library(ggthemes)
library(tinytex)
library(prettydoc) 
windowsFonts(Times = windowsFont("Times New Roman"))
library(ggthemes)
```


?datasets
library(help = "datasets")
?iris

iris
head(iris)

#basic scatterplot
```{r message=FALSE, warning=FALSE}
qplot(Petal.Width, Petal.Length, data=iris)
```

#basic scatterplot w/ species colored
```{r message=FALSE, warning=FALSE}
qplot(Petal.Width, Petal.Length, color=Species, data=iris)
```

#same thing as above, but strictly ggplot2
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species))+
      geom_point()
```

#size matches sepal length. jitter separates dots slightly for viewing. alpha makes dots more transparent
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Sepal.Length))+
  geom_jitter(alpha=0.5)
```

#size matches sepal length. jitter separates dots slightly for viewing. alpha makes dots more transparent. geom_smooth makes regression line with std. error
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Sepal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  theme(legend.position = "right")
```

#histogram
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Length, fill=Species))+
  geom_histogram()+
  theme(legend.position = "right")
```

#histogram w/ indiv. plots for species w/ facet_grid
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Length, fill=Species))+
  geom_histogram()+
  facet_grid(Species ~ .)
  theme(legend.position = "none")
```
  
#histogram w/ indiv. plots for species w/ facet_grid (geom_density is cleaner, less blocky)
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Length, fill=Species))+
  geom_density(alpha=0.5)+
  facet_grid(Species ~ .)
theme(legend.position = "none")
```

#facet_grid with previous scatterplots
```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Sepal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  facet_grid(Species ~ .)
  theme(legend.position = "right")
```
  
#Key Figures

windowsFonts(Times = windowsFont("Times New Roman"))
library(ggthemes)

```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Sepal.Width, Sepal.Length, color=Species, size=Sepal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  theme_fivethirtyeight(base_size = 12, base_family = "Times")+
  labs(title="Correlation Between Sepal Width and Sepal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  labs(x = "Sepal Width")+
  labs(y = "Sepal Length")+
  theme(legend.position = "right")
```

#SEPAL WIDTH VS. SEPAL LENGTH w/out theme
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Sepal.Width, Sepal.Length, color=Species, size=Sepal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  labs(title="Correlation Between Sepal Width and Sepal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  labs(x = "Sepal Width (cm)")+
  labs(y = "Sepal Length (cm)")+
  xlim(2.0, 4.5)+
  ylim(0.0, 10.0)+
  theme(legend.position = "bottom")
```

#SEPAL WIDTH VS. SEPAL LENGTH w/ theme
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Sepal.Width, Sepal.Length, color=Species, size=Sepal.Length))+
  geom_point(size=2)+ 
  geom_smooth(method=lm)+
  theme_fivethirtyeight(base_size = 12, base_family = "Times")+
  labs(title="Correlation Between Sepal Width and Sepal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  theme(axis.title = element_text()) + ylab('Sepal Length (cm)') + xlab('Sepal Width (cm)')+
  xlim(2.0, 4.5)+
  ylim(0.0, 10.0)+
  theme(legend.position = "bottom")
```



#PETAL WIDTH VS. PETAL LENGTH w/out theme
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Petal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  labs(title="Correlation Between Petal Width and Petal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  labs(x = "Petal Width (cm)")+
  labs(y = "Petal Length (cm)")+
  xlim(0.0, 3.0)+
  ylim(0.0, 8.0)+
  theme(legend.position = "bottom")
```

#PETAL WIDTH VS. PETAL LENGTH w/ theme & facet
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Petal.Length))+
  geom_point(size=2)+
  geom_smooth(method=lm)+
  labs(title="Correlation Between Petal Width and Petal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  labs(x = "Petal Width (cm)")+
  labs(y = "Petal Length (cm)")+
  xlim(0.0, 3.0)+
  ylim(0.0, 8.0)+
  facet_grid(Species ~ .)+
  theme(legend.position = "right")
```

#PETAL WIDTH VS. PETAL LENGTH w/ theme
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Petal.Length))+
  geom_point(size=2)+ 
  geom_smooth(method=lm)+
  theme_fivethirtyeight(base_size = 12, base_family = "Times")+
  labs(title="Correlation Between Petal Width and Petal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  theme(axis.title = element_text()) + ylab('Petal Length (cm)') + xlab('Petal Width (cm)')+
  xlim(0.0, 3.0)+
  ylim(0.0, 8.0)+
  theme(legend.position = "bottom")
```


#PETAL WIDTH VS. PETAL LENGTH w/ theme & facet
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(Petal.Width, Petal.Length, color=Species, size=Petal.Length))+
  geom_point(size=2)+ 
  geom_smooth(method=lm)+
  theme_fivethirtyeight(base_size = 12, base_family = "Times")+
  labs(title="Correlation Between Petal Width and Petal Length of Iris Species",
       caption="Data were collected by Anderson, Edgar (1935).")+
  theme(axis.title = element_text()) + ylab('Petal Length (cm)') + xlab('Petal Width (cm)')+
  xlim(0.0, 3.0)+
  ylim(0.0, 8.0)+
  facet_grid(Species ~ .)+
  theme(legend.position = "right")
```


#Boxplot NO ggplot2 Sepal Length
```{r message=FALSE, warning=FALSE}
boxplot(Sepal.Length~Species,
        data=iris, 
        main='Sepal Length by Species',
        sub='Data were collected by Anderson, Edgar (1935).',
        xlab='Species',
        ylab='Sepal Length (cm)',
        col='darkgreen',
        border='black')
```


#Boxplot w/ ggplot2 Sepal Length
```{r message=FALSE, warning=FALSE}
ggplot(iris, aes(x=Species, y=Sepal.Length, fill=Species)) + 
  geom_boxplot(width=0.5,lwd=1)+
  labs(title="Sepal Length by Species", x = "Species", y = "Sepal Length (cm)",  
       caption="Data were collected by Anderson, Edgar (1935).")
```


#Boxplot w/ ggplot2 Sepal Width
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(iris, aes(x=Species, y=Sepal.Width, fill=Species)) + 
  geom_boxplot(width=0.5,lwd=1)+
  labs(title="Sepal Length by Species", x = "Species", y = "Sepal Width (cm)",  
       caption="Data were collected by Anderson, Edgar (1935).")

```

#Boxplot w/ ggplot2 Petal Length
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(iris, aes(x=Species, y=Petal.Length, fill=Species)) + 
  geom_boxplot(width=0.5,lwd=1)+
  labs(title="Sepal Length by Species", x = "Species", y = "Petal Length (cm)",  
       caption="Data were collected by Anderson, Edgar (1935).")
```


#Boxplot w/ ggplot2 Petal Width
```{r message=FALSE, paged.print=FALSE}
ggplot(iris, aes(x=Species, y=Petal.Width, fill=Species)) + 
  geom_boxplot(width=0.5,lwd=1)+
  labs(title="Sepal Length by Species", x = "Species", y = "Petal Width (cm)",  
       caption="Data were collected by Anderson, Edgar (1935).")
```






























### AirPassengers 

**Loading Data / Basic Plotting Tutorial**
```{r message=FALSE, warning=FALSE}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(forecast) #for autoplot with geom_smooth linear regression
library(tseries) #for autoplot with geom_smooth linear regression
```

?datasets
library(help = "datasets")
?AirPassengers

AirPassengers
head(AirPassengers)

#this dataset does not have column names, but that won't stop us

class(AirPassengers) #ts is the output = this means the data is a time series 
start(AirPassengers) #shows first year 
end(AirPassengers) #shows last year


```{r message=FALSE, warning=FALSE}
plot(AirPassengers, xlab="Year", ylab="Air Passengers (1,000s)", main="Time Series of Air Passengers", 
     sub="The classic Box & Jenkins airline data. Monthly totals of international airline passengers, 1949 to 1960.")
```


```{r message=FALSE, warning=FALSE}
boxplot(AirPassengers~cycle(AirPassengers),xlab="Date", ylab = "Air Passengers (1,000s)" ,main ="Monthly Air Passengers Boxplot from 1949 to 1961")
```


```{r message=FALSE, warning=FALSE}
autoplot(AirPassengers) + 
  geom_smooth(method="lm")+ 
  labs(x ="Date", y = "Air Passengers (1,000s)", title="Air Passengers from 1949 to 1961") 
```
  
ARIMA is an acronym for “autoregressive integrated moving average.” It's a model used in statistics and econometrics to measure events that happen over a period of time.

```{r message=FALSE, warning=FALSE}
arimaAirPassengers <- auto.arima(AirPassengers)
arimaAirPassengers
```

If the residuals plot is around 0 w/ no movement, then ARIMA model is a good fit
```{r message=FALSE, warning=FALSE}
ggtsdiag(arimaAirPassengers)
```

95% confidence looking 48 months into the future
```{r message=FALSE, warning=FALSE}
forecastAirPassengers <- forecast(arimaAirPassengers, level = c(95), h = 48)
autoplot(forecastAirPassengers)
```




























### BJsales

**Loading Data / Basic Plotting Tutorial**
```{r message=FALSE, warning=FALSE}

library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(tseries) #for autoplot with geom_smooth linear regression
library(forecast) #for autoplot with geom_smooth linear regression
```

?datasets
library(help = "datasets")
?BJsales

BJsales
head(BJsales)

class(BJsales) #ts is the output = this means the data is a time series 
start(BJsales) #shows first year 
end(BJsales) #shows last year

```{r message=FALSE, warning=FALSE}
plot(BJsales, xlab="Time", ylab="Sales", main="Sales ", col.main="SteelBlue",
     sub="The data are given in Box & Jenkins (1976). Obtained from the Time Series Data Library at https://robjhyndman.com/TSDL/")
```


boxplot(BJsales~cycle(BJsales),xlab="Date (unknown scale)", ylab = "Sales (unknown scale)" ,main ="Sales Over Time")


```{r message=FALSE, warning=FALSE}
autoplot(BJsales) + 
  geom_smooth(method="lm")+ 
  labs(x ="Date (unknown scale)", y = "Sales (unknown scale)", title="Sales Over Time") 
```

#ARIMA is an acronym for “autoregressive integrated moving average.” It's a model used in statistics and econometrics to measure events that happen over a period of time.
```{r message=FALSE, warning=FALSE}
arimaBJsales <- auto.arima(BJsales)
arimaBJsales
```


#if the residuals plot is around - w/ no movement, then ARIMA model is a good fit
```{r message=FALSE, warning=FALSE}
ggtsdiag(arimaBJsales)
```


#95% confidence looking 10 months into the future
```{r message=FALSE, warning=FALSE}
forecastBJsales <- forecast(arimaBJsales, level = c(95), h = 10)
autoplot(forecastBJsales)
```



```{r message=FALSE, warning=FALSE}
forecastedJsales <- forecast(arimaBJsales, 10)
plot(forecastedJsales, main="Forecast of Sales Over Time", col.main="SteelBlue", ylab = "Sales (unknown scale)", xlab="Date (unknown scale)")
```





























### BOD  



**Loading Data / Basic Plotting Tutorial**
```{r message=FALSE, warning=FALSE}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
```

?datasets
library(help = "datasets")
?BOD

BOD
head(BOD)

class(BOD) #data frame

#same thing as basic ggplot
```{r message=FALSE, warning=FALSE}
qplot(Time, demand, data=BOD)
```

```{r message=FALSE, warning=FALSE}
ggplot(BOD, aes(Time, demand))+
  geom_point()
```


```{r message=FALSE, warning=FALSE}
ggplot(BOD, aes(Time, demand))+
  geom_line()+
  expand_limits(y=0, x=0)+
  labs(title="Biochemical Oxygen Demand Versus Time", x = "Time (days)", y = "Biochemical Oxygen Demand (mg/l)",  
       caption="Bates, D.M. and Watts, D.G. (1988), Nonlinear Regression Analysis and Its Applications, Wiley, Appendix A1.4.")+
  theme(plot.title = element_text(color = "SteelBlue"))

```































### CO2



**Loading Data / Basic Plotting Tutorial**
```{r message=FALSE, warning=FALSE}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
```


?datasets
library(help = "datasets")
?CO2

CO2
head(CO2)

class(CO2) #data frame

same thing as basic ggplot
```{r}
qplot(uptake, conc, data=CO2, fill=Type)
```




```{r}
ggplot(CO2, aes(x=Treatment, y=uptake)) + 
  geom_boxplot()
```

```{r}
t.test(CO2$uptake ~ CO2$Treatment)
```



```{r}
ggplot(CO2, aes(x=Type, y=uptake)) + 
  geom_boxplot()
```

```{r}
t.test(CO2$uptake ~ CO2$Type)
```


set the plotting area into a 1*2 array
```{r}
par(mfrow = c(1,2)) 
boxplot(CO2$uptake ~ CO2$Treatment)
boxplot(CO2$uptake ~ CO2$Type)
```


set the plotting area into a 1*1 array
```{r}
par(mfrow = c(1,1))
boxplot(CO2$uptake ~ CO2$Type + CO2$Treatment)
```


```{r}
par(mfrow = c(1,1))
boxplot(subset(CO2$uptake, CO2$conc == 95), 
        subset(CO2$uptake, CO2$conc == 175),
        subset(CO2$uptake, CO2$conc == 250),
        subset(CO2$uptake, CO2$conc == 350),
        subset(CO2$uptake, CO2$conc == 500),
        subset(CO2$uptake, CO2$conc == 675),
        subset(CO2$uptake, CO2$conc == 1000), 
        names = c("95", "175", "250", "350", "500", "675", "1000"))
```



```{r}
par(mfrow = c(2,2))
plot(uptake ~ conc, data = subset(CO2,Type == "Quebec" & Treatment == "chilled"), 
     main="Chilled Quebec", xlab = "Concentration (ML/L)", ylab = "Carbon Dioxide Uptake (μmol/m2s)" )
plot(uptake ~ conc, data = subset(CO2,Type == "Mississippi" & Treatment == "chilled"), 
     main="Chilled Mississippi", xlab = "Concentration (ML/L)", ylab = "Carbon Dioxide Uptake (μmol/m2s)")
plot(uptake ~ conc, data = subset(CO2,Type == "Quebec" & Treatment == "nonchilled"),
     main="Nonchilled Quebec", xlab = "Concentration (ML/L)", ylab = "Carbon Dioxide Uptake (μmol/m2s)")
plot(uptake ~ conc, data = subset(CO2,Type == "Mississippi" & Treatment == "nonchilled"), 
     main="Nonchilled Mississippi", xlab = "Concentration (ML/L)", ylab = "Carbon Dioxide Uptake (μmol/m2s)")
```



```{r}
ggplot(CO2, aes(conc, uptake, color=Treatment, shape=Type)) + 
  geom_point()+
  expand_limits(y=0, x=0)+
  labs(title="CO2 Uptake of Plants from Mississippi and Quebec", x = "Concentration (ML/L)", y = "Carbon Dioxide Uptake (μmol/m2s)",  
       caption="The CO2 data frame has 84 rows and 5 columns of data from an experiment on the cold tolerance of the grass species Echinochloa crus-galli.")
```


```{r}
ggplot(CO2, aes(conc, uptake, color=Plant, type=Treatment)) + 
  geom_line()+
  geom_point()+
  expand_limits(y=0, x=0)+
  labs(title="CO2 Uptake of Plants from Mississippi and Quebec", x = "Concentration (ML/L)", y = "Carbon Dioxide Uptake (μmol/m2s)",  
       caption="The CO2 data frame has 84 rows and 5 columns of data from an experiment on the cold tolerance of the grass species Echinochloa crus-galli.")
```






























### ChickWeight



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
```

?datasets
library(help = "datasets")
?ChickWeight

ChickWeight
head(ChickWeight)

class(ChickWeight) #data frame

The body weights of the chicks were measured at birth and every second day thereafter until day 20. 
They were also measured on day 21. There were four groups on chicks on different protein diets.

tail(ChickWeight)

```{r}
chick0 <- ChickWeight[ChickWeight$Time %in% 0:5, ]
chick1 <- ChickWeight[ChickWeight$Time %in% 6:10, ]
chick2 <- ChickWeight[ChickWeight$Time %in% 11:15, ]
chick3 <- ChickWeight[ChickWeight$Time %in% 16:21, ]
```


chick0
chick1
chick2
chick3



```{r}
par(mfrow = c(2,2)) 
boxplot(chick0$weight ~ chick0$Diet, 
        main = "Chick Weight By Diet Type (0-5 Days)",
        xlab = "Diet Types",
        ylab = "Weight",
        ylim = c(0, 300))
boxplot(chick1$weight ~ chick1$Diet, 
        main = "Chick Weight By Diet Type (6-10 Days)",
        xlab = "Diet Types",
        ylab = "Weight",
        ylim = c(0, 300))
boxplot(chick2$weight ~ chick2$Diet, 
        main = "Chick Weight By Diet Type (11-15 Days)",
        xlab = "Diet Types",
        ylab = "Weight",
        ylim = c(0, 300))
boxplot(chick3$weight ~ chick3$Diet, 
        main = "Chick Weight By Diet Type (16-21 Days)",
        xlab = "Diet Types",
        ylab = "Weight",
        ylim = c(0, 300))
```






```{r}
chickDiet1 <- ChickWeight[ChickWeight$Diet == 1, ]
chickDiet2 <- ChickWeight[ChickWeight$Diet == 2, ]
chickDiet3 <- ChickWeight[ChickWeight$Diet == 3, ]
chickDiet4 <- ChickWeight[ChickWeight$Diet == 4, ]
```


chickDiet1
chickDiet2
chickDiet3
chickDiet4



```{r}
par(mfrow = c(2,2)) 
boxplot(chickDiet1$weight ~ chickDiet1$Time, 
        main = "Chick Weight By Diet Type (1)",
        xlab = "Days",
        ylab = "Weight (gm)",
        ylim = c(0, 300))
boxplot(chickDiet2$weight ~ chickDiet2$Time, 
        main = "Chick Weight By Diet Type (2)",
        xlab = "Days",
        ylab = "Weight (gm)",
        ylim = c(0, 300))
boxplot(chickDiet3$weight ~ chickDiet3$Time, 
        main = "Chick Weight By Diet Type (3)",
        xlab = "Days",
        ylab = "Weight (gm)",
        ylim = c(0, 300))
boxplot(chickDiet4$weight ~ chickDiet4$Time, 
        main = "Chick Weight By Diet Type (4)",
        xlab = "Days",
        ylab = "Weight (gm)",
        ylim = c(0, 300))

```





same thing as basic ggplot
```{r}
qplot(Time, weight, data=ChickWeight, color=factor(Diet), facets = Diet ~ .,
      geom=c("point", "smooth"), 
      main="Protein Diet Versus Chick Weight ", xlab = "Time (days)", ylab = "Chick Weight (gm)" )
```



```{r}
ggplot(ChickWeight, aes(x=Diet, y=weight)) + 
  geom_boxplot()
```



































### DNase



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
```

?datasets
library(help = "datasets")
?DNase

The DNase data frame has 176 rows and 3 columns of data obtained during development of 
an ELISA assay for the recombinant protein DNase in rat serum.

DNase
head(DNase)

class(DNase) #data frame

Looking at distribution of data 
```{r}
par(mfrow = c(1,1))
hist(DNase$density, breaks=10, main = "")
```

```{r}
boxplot(density ~ Run, data = DNase)
```



Plotting outliers
```{r}
par(mfrow = c(1,1))
boxplot(DNase$density~DNase$conc,xlab="Concentration of Protein (ng/ml)",ylab="Optical Density")
```

Finding outliers
```{r}
boxplot.stats(DNase$density[DNase$conc==12.5])$out
boxplot.stats(DNase$density[DNase$conc==6.25])$out
```

Removing outliers
```{r}
df_no_outliers <- subset(x=DNase, !density %in% boxplot.stats(DNase$density[DNase$conc==12.5])$out & !density %in% boxplot.stats(DNase$density[DNase$conc==6.25])$out ) 

```

Final Graph
```{r}
ggplot(df_no_outliers, aes(x=conc, y=density)) + 
  geom_smooth()+
  geom_point()+
  expand_limits(y=0, x=0)+
  labs(title="Recombinant Protein DNase ELISA Assay Density Versus Concentration", x = "Concentration of Protein (ng/ml)", y = "Optical Density",  
       caption="The DNase data frame has 176 rows and 3 columns of data obtained during development of an ELISA assay for the recombinant protein DNase in rat serum.")
```































### EUStockMarkets



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(broom)
library(dplyr)

```

?datasets
library(help = "datasets")
?EuStockMarkets

Contains the daily closing prices of major European stock indices: Germany DAX (Ibis), 
Switzerland SMI, France CAC, and UK FTSE. The data are sampled in business time, i.e., weekends and holidays are omitted.

EuStockMarkets
head(EuStockMarkets)

class(EuStockMarkets) #timeseries

```{r}
EuStockDF<-as.data.frame(EuStockMarkets) 
EuStockDF$Date <- as.numeric(time(EuStockMarkets)) 
```

now the dates have an official column
head(EuStockDF) 

```{r}
ggplot(EuStockDF,aes(x = Date))+
  expand_limits(y=c(2000, 9000))+
  labs(title="European Stock Indices 1991 to 1998", x = "Year", y = "Stock Price",  
       caption="Contains the daily closing prices of major European stock indices: Germany DAX (Ibis), Switzerland SMI,               France CAC,and UK FTSE. The data are sampled in business time, i.e., weekends and holidays are omitted.")+
  geom_line(aes(y = DAX), color="darkslateblue")+
  geom_line(aes(y = SMI), color="darkorchid")+
  geom_line(aes(y = CAC), color="mediumseagreen")+
  geom_line(aes(y = FTSE), color="indianred3")
```


```{r}
tidyEuStockMarkets <- tidy(EuStockMarkets)
tidyEuStockMarkets <- tidyEuStockMarkets %>%
  rename(Date=index,Stock_Index=series,Price=value)
```

```{r}
tidyEuStockMarkets
```


```{r}
ggplot(tidyEuStockMarkets,aes(x=Date,y=Price))+
  geom_line(aes(color=Stock_Index))+
  expand_limits(y=c(2000, 10000))+
  labs(title="European Stock Indices 1991 to 1998", x = "Year", y = "Stock Price",  
       caption="Contains the daily closing prices of major European stock indices: Germany DAX (Ibis), Switzerland SMI,           France CAC,and UK FTSE. The data are sampled in business time, i.e., weekends and holidays are omitted.")
```































### HairEyeColor



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
```

?datasets
library(help = "datasets")
?HairEyeColor

#Distribution of hair and eye color and sex in 592 statistics students.

HairEyeColor
head(HairEyeColor)

class(HairEyeColor) #table
  
```{r}
ggplot(HairEyeColor, aes(Eye, Freq, fill=Sex))+
    geom_boxplot()
```

```{r}
ggplot(HairEyeColor, aes(Hair, Freq, fill=Sex))+
  geom_boxplot()
```

```{r}
ggplot(HairEyeColor, aes(Hair, fill=Hair))+
  scale_fill_manual(values=c("#1C1504", "#5E4303", "#F07743", "#FFFBF1"))+
  geom_density(alpha=0.7)

```

```{r}
ggplot(HairEyeColor, aes(Eye, fill=Eye,))+
  scale_fill_manual(values=c("#3B1002", "cyan3", "#93820B", "chartreuse4"))+
  geom_density(alpha=0.7)

```








```{r}
library(vcd)
mosaic(HairEyeColor, shade = TRUE)
```
  






```{r}
HairEyeColorchi <- HairEyeColor[,,"Male"] + HairEyeColor[,,"Female"]
HairEyeColorchi
#p-value < 2.2e-16 significant
chisq.test(HairEyeColorchi)
```

HairEyeColorchi/sum

```{r}
barplot(HairEyeColorchi, beside=TRUE, legend.text=TRUE, xlab="Eye Color", ylab="Frequency", main="Eye Color vs. Hair Color", 
        col=c("#1C1504", "#5E4303", "#F07743", "#FFFBF1")) 
```






HairEyeColor
```{r}
#2 means columns, 1 means rows
MH <- apply(HairEyeColor[,,"Male"],1,sum)
MH
FH <- apply(HairEyeColor[,,"Female"],1,sum)
FH

HairEyeColorchi2 <- cbind(MH,FH)
HairEyeColorchi2
#p-value = 0.04613 somehwta significant 
chisq.test(HairEyeColorchi2)
```

```{r}
barplot(HairEyeColorchi2, beside=TRUE, legend.text=TRUE, xlab="Gender", ylab="Frequency", main="Gender vs. Hair Color", 
        names.arg=c("Male","Female"), col=c("#1C1504", "#5E4303", "#F07743", "#FFFBF1"))
```









HairEyeColor

```{r}
#2 means columns, 1 means rows
ME <- apply(HairEyeColor[,,"Male"],2,sum)
FE <- apply(HairEyeColor[,,"Female"],2,sum)
HairEyeColorchi3 <- cbind(ME,FE)

HairEyeColorchi3
#p-value = 0.6754 not significant 
chisq.test(HairEyeColorchi3)
```

```{r}
barplot(HairEyeColorchi3, beside=TRUE, legend.text=TRUE, xlab="Gender", ylab="Frequency", main="Gender vs. Eye Color",
        names.arg=c("Male","Female"), col=c("#3B1002", "cyan3", "#93820B", "chartreuse4"))
```



































### Harman23.cor



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(ggcorrplot)
```

?datasets
library(help = "datasets")
?Harman23.cor

A correlation matrix of eight physical measurements on 305 girls between ages seven and seventeen.
```{r}
Harman23.cor
head(Harman23.cor)

class(Harman23.cor) #list
```

Exporting file to csv so I can manipulate it to work more easily with corrplot
getwd()
write.csv(Harman23.cor,"C:/Users/Siggi/Downloads/Harman23.cor.csv", row.names = FALSE)

```{r}
Harman23DF <- as.data.frame(Harman23.cor,  col.names = c("", "", ""))#dataframe now
Harman23DF #dataframe now so ggcorrplot will work
class(Harman23DF)
```

Anything correlated to itself will be 1 so ignore middle diagonal and anything paired to itself
```{r}
ggcorrplot(Harman23DF, hc.order = TRUE, outline.col = "white")
```








```{r}
library(corrplot)
```

```{r}
#removing unnecessary columns
Harman23DF$c.0..0..0..0..0..0..0..0. <- NULL
Harman23DF$X305 <- NULL
Harman23DF
```

```{r}
corrplot(as.matrix(Harman23DF), is.corr = TRUE, method = "square")
```

```{r}
corrplot(as.matrix(Harman23DF), is.corr = TRUE, method = "square", diag = FALSE)
```






```{r}
testHarman23DF = cor.mtest(Harman23DF, conf.level = 0.95)
testHarman23DF
#could not get this to work perfectly
corrplot(as.matrix(Harman23DF), is.corr = TRUE, method = "square", diag = FALSE, p.mat = testHarman23DF$p, sig.level = 0.05)
```
 

Main problem was reformatting original dataset to work with the corrplot function







































### mtcars



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(ggcorrplot)
```


?datasets
library(help = "datasets")
?mtcars

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel 
consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

mtcars
head(mtcars)

Only would do this if I want to export file to csv so I can manipulate it to work more easily with corrplot / renaming
getwd()
write.csv(mtcars,"C:/Users/Siggi/Downloads/mtcars.csv", row.names = FALSE)

renaming columns 
```{r}
mtcarsnew <- rename(mtcars, "# Gears" = gear, "# Carburetors" = carb, Transmission = am, 
                    Engine = vs, "1/4 Mile Time" = qsec, Weight = wt, "Rear Axle Ratio" = drat, 
                    Horsepower = hp, Displacement = disp, "# Cylinders" = cyl, MPG = mpg)
```

mtcarsnew

```{r}
class(mtcarsnew) #data frame
mtcarsCOR = cor(mtcarsnew)
mtcarsCOR
```


```{r}
ggcorrplot(mtcarsCOR, hc.order = TRUE, outline.col = "white")

```

```{r}
corrplot(mtcarsCOR, is.corr = TRUE, method = "square", diag = FALSE, order = 'AOE', addCoef.col = "gray")
```


```{r}
testmtcarsCOR = cor.mtest(mtcarsCOR, conf.level = 0.95)
testmtcarsCOR
```

```{r}
corrplot(mtcarsCOR, is.corr = TRUE, method = "square", diag = FALSE, order = 'AOE', addCoef.col = "gray", p.mat = testmtcarsCOR$p, sig.level = 0.05)
```







































### Indometh



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(ggcorrplot)
```

?datasets
library(help = "datasets")
?Indometh

The Indometh data frame has 66 rows and 3 columns of data on the pharmacokinetics of indometacin (or, older spelling, ‘indomethacin’).

Indometh
head(Indometh)

If I want to view raw csv
write.csv(Indometh,"C:/Users/Siggi/Downloads/Indometh.csv", row.names = FALSE)

class(Indometh) #grouped data AND data frame

```{r}
boxplot(time ~ Subject, data = Indometh)
boxplot(conc ~ Subject, data = Indometh)
```

```{r}
ggplot(Indometh, aes(x=time, y=conc, color=Subject)) + 
  geom_smooth()+
  geom_point()+
  labs(title="Indometacin Plasma Concentration vs. Time", x = "Time (hr)", y = "Plasma Concentrations of Indometacin (mcg/ml)",  
       caption="The Indometh data frame has 66 rows and 3 columns of data on the pharmacokinetics of indometacin")
```


```{r message=FALSE, warning=FALSE}
library(plotly)
```



```{r}
plot_ly(data = Indometh, x=~time, y=~conc, type = 'scatter', mode = 'lines', color=~Subject) %>% 
  arrange(Subject) %>% 
  layout(title="Indometacin Plasma Concentration vs. Time", 
         xaxis = list(title = "Time (hr)"), 
         yaxis = list(title = "Plasma Concentrations of Indometacin (mcg/ml)"),
         legend=list(title=list(text='<b> Subject </b>')),
         annotations = list(x = 1.0, y = -0.1, showarrow = F, xref='paper', yref='paper', 
         xanchor='right', yanchor='auto', xshift=0, yshift=0, text = "The Indometh data frame has 66 rows and 3 columns of             data on the pharmacokinetics of indometacin"))
```

































### InsectSprays



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(ggcorrplot)
```

?datasets
library(help = "datasets")
?InsectSprays

The counts of insects in agricultural experimental units treated with different insecticides.

InsectSprays
head(InsectSprays)

If I want to view raw csv
write.csv(InsectSprays,"C:/Users/Siggi/Downloads/InsectSprays.csv", row.names = FALSE)

class(InsectSprays) #data frame




```{r}
#Either method for boxplot 1.)
ggplot(InsectSprays, aes(x=spray, y=count))+
  geom_boxplot()
```

```{r}
#2.)
boxplot(count ~ spray, data = InsectSprays,
        xlab = "Type of spray", ylab = "Insect count",
        main = "InsectSprays data", varwidth = TRUE, col = "lightgray")
```

  
```{r}
ANOVAsprays <- aov(count ~ spray, data = InsectSprays)
```

```{r}
summary(ANOVAsprays) #Reject null hypothesis. P-value less than 0.05. There is a sig difference.
anova(ANOVAsprays) #same thing as using summary
```

```{r}
pairwise.t.test(x=InsectSprays$count, g=InsectSprays$spray, p.adj="bonf")
```

```{r}
plot(ANOVAsprays)
```







```{r}
#Using SQRT to satisfy ANOVA conditions and ensure results are more accurate
ggplot(InsectSprays, aes(x = spray, y = sqrt(count))) + 
  geom_boxplot()
```

```{r}
boxplot(sqrt(count) ~ spray, data = InsectSprays,
        xlab = "Type of spray", ylab = "Insect count",
        main = "InsectSprays data", varwidth = TRUE, col = "lightgray")
```


```{r}
ANOVAspraysSQRT <- aov(sqrt(count) ~ spray, data = InsectSprays)
```

```{r}
summary(ANOVAspraysSQRT)
anova(ANOVAspraysSQRT)
```

```{r}
plot(ANOVAspraysSQRT)
```



```{r}
t.test(InsectSprays$count[InsectSprays$spray == 'A'], InsectSprays$count[InsectSprays$spray == 'B']) #not significant
t.test(InsectSprays$count[InsectSprays$spray == 'A'], InsectSprays$count[InsectSprays$spray == 'C']) #significant
t.test(InsectSprays$count[InsectSprays$spray == 'A'], InsectSprays$count[InsectSprays$spray == 'D']) #significant
t.test(InsectSprays$count[InsectSprays$spray == 'A'], InsectSprays$count[InsectSprays$spray == 'E']) #significant
t.test(InsectSprays$count[InsectSprays$spray == 'A'], InsectSprays$count[InsectSprays$spray == 'F']) #not significant
```














































### JohnsonJohnson                                        



**Loading Data / Basic Plotting Tutorial**

```{r}
library(ggplot2) #qplot (basic scatterplot) 
library(ggthemes)
library(tidyverse)
library(datasets)
library(ggfortify)
library(rmarkdown)
library(ggcorrplot)
library(tseries) #for autoplot with geom_smooth linear regression
library(forecast)
```

?datasets
library(help = "datasets")
?JohnsonJohnson

Quarterly earnings (dollars) per Johnson & Johnson share 1960–80.

Sources:
https://rpubs.com/scarden/757860
https://towardsdatascience.com/time-series-forecasting-in-r-with-holt-winters-16ef9ebdb6c0
https://www.statology.org/ljung-box-test/
https://www.r-bloggers.com/2012/07/holt-winters-forecast-using-ggplot2/ 

JohnsonJohnson
head(JohnsonJohnson)
str(JohnsonJohnson)

If I want to view raw csv
write.csv(JohnsonJohnson,"C:/Users/Siggi/Downloads/JohnsonJohnson.csv", row.names = FALSE)

class(JohnsonJohnson) #timeseries


```{r}
plot(JohnsonJohnson, ylab = 'Earnings', main = 'Quarterly Earnings of Johnson & Johnson')
#the log will reduce variance
plot(log(JohnsonJohnson), ylab = 'Log of Earnings', main = 'Log Transformed Quarterly Earnings of Johnson & Johnson') 
```


```{r}
#If I want to use ggplot you have to switch to data frame and make Date numeric column
JohnsonJohnsonDF<-as.data.frame(JohnsonJohnson) 
JohnsonJohnsonDF$Date <- as.numeric(time(JohnsonJohnson)) 
JohnsonJohnsonDFlog <- log(JohnsonJohnsonDF$x)
JohnsonJohnsonDFlog

ggplot(JohnsonJohnsonDF, aes(x=Date, y=JohnsonJohnson))+
  geom_line()
ggplot(JohnsonJohnsonDF, aes(x=Date, y=JohnsonJohnsonDFlog))+
  geom_line()
```




```{r}
plot(decompose(log(JohnsonJohnson)))
```


```{r}
JJHW1 <- HoltWinters(JohnsonJohnson)
JJHW1
```

```{r}
#Custom HoltWinters fitting
JJHW2 <- HoltWinters(JohnsonJohnson, alpha=0.2, beta=0.1, gamma=0.1)
JJHW2

#Visually evaluate the fits
plot(JohnsonJohnson, ylab="value", xlim=c(1960,1981))
lines(JJHW1$fitted[,1], lty=2, col="blue")
lines(JJHW2$fitted[,1], lty=2, col="red")
```


```{r}
#Predicting 24 months into future with 95% confidence. 
JJHW1.pred <- predict(JJHW1, 24, prediction.interval = TRUE, level=0.95)
#Visually evaluate the prediction
plot(JohnsonJohnson, ylab="value", xlim=c(1960,1983), ylim=c(0,25))
lines(JJHW1$fitted[,1], lty=2, col="blue")
lines(JJHW1.pred[,1], col="red")
lines(JJHW1.pred[,2], lty=2, col="brown")
lines(JJHW1.pred[,3], lty=2, col="brown")
```


```{r}
plot(JJHW1, JJHW1.pred)
```




```{r}
#Seasonality prediction 
JJHW3 <- HoltWinters(JohnsonJohnson, seasonal = "multiplicative")
JJHW3.pred <- predict(JJHW3, 24, prediction.interval = TRUE, level=0.95)
plot(JohnsonJohnson, ylab="value", xlim=c(1960,1983), ylim=c(0,25))
lines(JJHW3$fitted[,1], lty=2, col="blue")
lines(JJHW3.pred[,1], col="red")
lines(JJHW3.pred[,2], lty=2, col="brown")
lines(JJHW3.pred[,3], lty=2, col="brown")
```



```{r}
#Using forecast (similar to ARIMA)
JJHW1_forecast <- forecast(JJHW1, h=12, level=c(80,95))
plot(JJHW1_forecast, xlim=c(1960,1983))
lines(JJHW1_forecast$fitted, lty=2, col="red")
```


#acf bars should be within blue bars if there is correlation of fit residuals.
acf(JJHW1_forecast$residuals, lag.max=20, na.action=na.pass)
#Ideally, we would like to fail to reject the null hypothesis. That is, 
#we would like to see the p-value of the test be greater than 0.05 because
#this means the residuals for our time series model are independent, which 
#is often an assumption we make when creating a model.
Box.test(JJHW1_forecast$residuals, lag=20, type="Ljung-Box")
#no skew is good!
hist(JJHW1_forecast$residuals)
```








ARIMA METHOD

```{r}
#ARIMA is an acronym for “autoregressive integrated moving average.” It's a model used in statistics and econometrics to measure events that happen over a period of time.
arimaJohnsonJohnson <- auto.arima(JohnsonJohnson)
arimaJohnsonJohnson

#if the residuals plot is around - w/ no movement, then ARIMA model is a good fit
ggtsdiag(arimaJohnsonJohnson)
       
#95% confidence looking 10 months into the future
forecastarimaJohnsonJohnson <- forecast(arimaJohnsonJohnson, level = c(80,95), h = 12)
autoplot(forecastarimaJohnsonJohnson)
```

